---
title: "Explaining the Restaurant Inspection Dashboard"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

The data used on this [dashboard](./dashboard.html) is sourced from The New York City Department of Health and Mental Hygiene (DOHMH) food inspection data. I accessed the dataset, data dictionary, and other information through the P8105 course [website](https://p8105.com/dataset_restaurant_inspections.html).

I filtered the original `nyc_inspec` dataset to remove boroughs that are marked as "Missing" and, in the spirit of Halloween, the spookiest inspection grades: P (Pending on re-opening after an initial inspection resulting in a closure), Z (Pending), and Not Yet Graded (just to see how this compares to P and Z).

## Data Cleaning

See below for my data cleaning code.
```{r data cleaning, warning=FALSE, message=FALSE, results='hide'}
library(flexdashboard)
library(tidyverse)
library(viridis)
library(p8105.datasets)
library(plotly)

theme_set(theme_bw() + theme(legend.position = "bottom"))

data(rest_inspec)

inspect = rest_inspec %>% 
  janitor::clean_names() %>% 
  filter(!boro %in% c("Missing"),
         grade %in%  c("P", "Z", "Not Yet Graded")) %>% 
  mutate(grade = recode(grade, "P" = "P: Pending on re-opening after an initial inspection resulting in a closure", "Z" = "Z: Pending")) %>% 
  select(boro, grade, score, critical_flag, inspection_date, cuisine_description)
```


## Chart A
#### Distribution of Scores among Spookiest (Pending) Grades

To make this boxplot, I ordered the grades by score and plotted borough against score to see distribution of inspection scores within the lowest grades/those not yet graded. I grouped these by borough as well so I could visualize differences in scores/grades across boroughs.

Code for Chart A:

```{r chart A, warning=FALSE, message=FALSE, results='hide'}
inspect %>% 
  mutate(grade = fct_reorder(grade, score)) %>% 
  plot_ly(x = ~boro,
          y = ~score,
          color = ~grade,
          type = "box",
          colors = viridis(3, begin = 0.2, end = 0.8, option = "A")) %>% 
  layout(xaxis = list(title = "Borough"),
         yaxis = list(title = "Inspection Score"),
         boxmode = "group",
         legend = list(orientation = 'h'))
```

## Chart B
#### Grades Among "Critical" Restaurants (hauntingly high!)

To keep up with the spooky theme, I created a bar chart including grades only among "Critical" restaurants - meaning those that are most likely to contribute to foodborne illness. For example, we can observe that there is a far higher count of `A` grades than `P` or `Z` among "Critical" restaurants.

Code for Chart B:

```{r chart B, warning=FALSE, message=FALSE, results='hide'}
rest_inspec %>% 
  filter(critical_flag == "Critical") %>%  
  count(grade) %>%
  mutate(grade = fct_reorder(grade, n)) %>% #to make bar chart in a particular order
  plot_ly(x = ~n,
          y = ~grade,
          color = ~grade,
          type = "bar",
          colors = viridis(6, begin = 0.1, end = 0.9, option = "A")) %>% 
  layout(xaxis = list(title = "Count"), yaxis = list(title = "Grade"))
```

## Chart C
#### Pending/Not Graded Inspection Scores among Pizza and Pizza/Italian Restaurants by Date

Among cuisines serving `Pizza` or `Pizza/Italian`, I created a scatterplot showing scores by date of inspection, coloring the points based on the grade. 

Code for Chart C:

```{r chart C, warning=FALSE, message=FALSE, results='hide'}
inspect %>% 
  filter(cuisine_description %in% c("Pizza", "Pizza/Italian")) %>% 
  plot_ly(x = ~inspection_date,
          y = ~score,
          color = ~grade,
          type = "scatter",
          colors = viridis(3, begin = 0.3, end = 0.9, option = "A")) %>% 
  layout(xaxis = list(title = "Date of Inspection"),
         yaxis = list(title = "Inspection Score"),
         legend = list(x = 0, y = 1))
```